The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-10-09 article.tex
More minor tweaks

Substructure matching is fundamental cheminformatics operation and
plays a key role in many other functions such as fingerprint,
descriptor generation and atom typing. Since CDK 1.2, functionality
has been added to handle the SMARTS query language. The SMARTS
language is supported well including advanced features such as
stereochemistry, component grouping, and atom
maps (to match reaction transformations). A new \textit{Pattern} API
has been added to CDK \cdkversion{}, which simplifies finding,
filtering, and transform results. The API is immutable allowing a
pattern to be initialized once and then matched against several
molecules or reactions across multiple threads. During initialization
the pattern is inspected as to determine what invariants will be
needed (e.g. ring size) and only calculating what is needed. The
internal matching algorithms provide a lazy iterator, such that the
next match is only computed when it is needed. The API handles
reactions in addition to molecules, and both can be specified as
either queries or targets.

@@

Ring finding is a key functionality in a cheminformatics library, and
the CDK knows a long history of ring finding~\cite{May2014}. In
particular, non-redundant ring sets have seen particular interest,
such as the smallest set of smallest rings, for which the CDK
implements two classical algorithms~\cite{Figueras1996,Berger2004}.
Recent work has implemented a new, faster algorithm, allowing
searching for various types of (non-redundant) ring
sets~\cite{May2014}. These are available via the new Cycles API:

>>>>

Substructure matching is fundamental cheminformatics operation and
plays a key role in many other functions such as fingerprint and
descriptor generation, and atom typing. Since CDK 1.2, functionality
has been added to handle the SMARTS query language. The SMARTS
language is supported well including advanced features such as
stereochemistry, component grouping, and atom
maps (to match reaction transformations). A new \textit{Pattern} API
has been added to CDK \cdkversion{}, which simplifies finding,
filtering, and transforming search results. The API is immutable allowing a
pattern to be initialized once and then matched against several
molecules or reactions across multiple threads. During initialization
the pattern is inspected as to determine what invariants will be
needed (e.g. ring size) and only calculating what is needed. The
internal matching algorithms provide a lazy iterator, such that the
next match is only computed when it is needed. The API handles
reactions in addition to molecules, and both can be specified as
either queries or targets.

@@

Ring finding is another key functionality in a cheminformatics library, and
the CDK knows a long history of ring finding~\cite{Berger2004,May2014}. Specifically,
non-redundant ring sets have seen particular interest,
such as the smallest set of smallest rings, for which the CDK
implements two classical algorithms~\cite{Figueras1996,Berger2004}.
Recent work has implemented a new, faster algorithm, allowing
searching for various types of (non-redundant) ring
sets~\cite{May2014}. These are available via the new Cycles API:
